---
title: "Fuel Efficiency and Transmission"
author: 'Aman Jindal'
date: '29 Aug 2020'
output:
  html_document:
    keep_md: yes
  word_document: default
  pdf_document: default
---

```{r, label='setup', echo=FALSE, include=FALSE}
knitr::opts_chunk$set(comment='', message=FALSE, warning=FALSE)
library(data.table)
library(ggplot2)
library(datasets)
library(gridExtra)
```

### 1. Executive Summary:

This report analyzes the relationship between fuel efficiency and transmission type for cars using the [mtcars](https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/mtcars) dataset. The following questions are answered:

i) Is an automatic or manual transmission better for MPG
ii) Quantify the MPG difference between automatic and manual transmissions

**First**, exploratory data analysis has been done to identify the relationship between `mpg` (fuel efficiency) and `am` (transmission) and confounding variables affecting the relationship.
**Second**, several multivariable linear regression models are run to identify the model that best answers our questions.
**Third**, residual diagnostics are run on the chosen model. <br>

Our analysis has revealed that **transmission type has no effect on MPG** and thus there is no `mpg` difference between automatic and manual transmissions.

### 2. Exploratory Data Analysis (EDA):

**i) Steps**: **First**, we make the variables `am` and `vs` (engine type) as factor variables. **Second**, we summarize the dataset to view the summary statistics (Appendix A). **Third**, we plot `mpg` against `am` (Appendix B), and `mpg` against `am` controlling for `cyl`. **Fourth**, we plot `mpg` against other variables and color the points according to `am` (Appendix C) <br>

**ii) Inferences**: At first glance, Manual transmission seems to have a better `mpg`. However, after examining the effect of other variables (`cyl`, `disp`, `hp`, `wt`, `drat`), the relationship seems to vanish.

```{r, results='hold', fig.height = 2.5, fig.width = 5}
data("mtcars"); setDT(mtcars)
mtcars[,`:=`(am = factor(am, labels=c('automatic','manual')), 
             vs = factor(vs, labels=c('vshaped','straight')))]
ggplot(data=mtcars) + geom_boxplot(mapping = aes(x=am, y=mpg)) + facet_grid(~cyl)
```


### 3. Model Fitting and Selection:

**i) Steps**: We start by fitting a linear model between `mpg` and `am` and then we add other variables that were diagnosed as confounders in EDA. We first add `wt`, then we add `cyl` (`disp`, `hp` and `cyl` are similar variables for cars. Hence, we have taken one of these). Then, we add `drat` and `vs`. We then perform analysis of variance (ANOVA) for our five models. <br>

**ii) Inferences**: Based on the results of ANOVA, we choose Model 3 as it is statistically significant at 0.01% level and has low Residual sum of squares (RSS). 

```{r}
fit1 <- lm(mpg~am, data=mtcars) 
fit2 <- lm(mpg~am+wt, data=mtcars)
fit3 <- lm(mpg~am+wt+cyl, data=mtcars) 
fit4 <- lm(mpg~am+wt+cyl+drat, data=mtcars)
fit5 <- lm(mpg~am+wt+cyl+drat+vs, data=mtcars)
anova(fit1, fit2, fit3, fit4, fit5)[,1:6]
```

### 4. Analyzing the chosen Model:

**i) ** `am` **variable**: The model reveals a p value for 0.89 for `am` variable. Also, the 95% confidence interval for `am` includes 0. <br>

**ii) Residuals diagnostics (Appendix D)**: The distribution of residuals is random about the fitted values. Residuals are i.i.d. and are normally distributed as evident from the Q-Q plot.
    
```{r}
summary(fit3)$coefficients 
confint.lm(fit3) # par(mfrow = c(2,2)); plot(fit3)
```

### 5. Conclusion:

i) There is not enough evidence at 5% significance level to conclude that manual transmission's `mpg` is different from that of automatic transmission. 
ii) Owing to the result above, we cannot conclude that at 5% significance level, the difference in `mpg` for two transmission types is different from 0.

### Appendix A: Summary Statistics of Variables

```{r}
summary(mtcars)
```

### Appendix B: `mpg` vs `am` boxplot

```{r, fig.height=4, fig.width=4}
g <- ggplot(data=mtcars)
g + geom_boxplot(mapping = aes(x=am, y=mpg))
```

### Appendix C: `mpg` vs other variables scatterplot (colour coded by `am`)

```{r, fig.height=12, fig.width=12}
g1 <- g + geom_point(mapping = aes(x=wt, y=mpg, col=am)) + theme_gray(base_size = 16) + theme(legend.position = 'none')
g2 <- g + geom_point(mapping = aes(x=cyl, y=mpg, col=am)) + theme_gray(base_size = 16) + theme(legend.position = 'none')
g3 <- g + geom_point(mapping = aes(x=disp, y=mpg, col=am)) + theme_gray(base_size = 16) + theme(legend.position = 'none')
g4 <- g + geom_point(mapping = aes(x=hp, y=mpg, col=am)) + theme_gray(base_size = 16) + theme(legend.position = 'none')
g5 <- g + geom_point(mapping = aes(x=drat, y=mpg, col=am)) + theme_gray(base_size = 16) + theme(legend.position = 'bottom')
g6 <- g + geom_point(mapping = aes(x=vs, y=mpg, col=am)) + theme_gray(base_size = 16) + theme(legend.position = 'bottom')

grid.arrange(g1, g2, g3, g4, g5, g6, nrow=3, ncol=2)
```

### Appendix D: Residual Diagnostics

```{r, figsize=12, figsize=14}
par(mfrow = c(2,2)); plot(fit3)
```




